---
title: "R Code Guide"
subtitle: "Case File #1997-US-S"
date: "September 4, 2025"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# Turn off evaluation for all chunks
knitr::opts_chunk$set(eval = FALSE)
```

## Getting Started

```{r}
# First, load your dataset 
data <- read.csv("1997_data.csv")
```

## Phase 1: Understanding Your Dataset's Size and Structure

### How big is your dataset?

```{r}
# Get the dimensions (rows x columns)
dim(data)
```

**INTEL** 

**What this does**: Shows you [number of rows, number of columns]. Rows usually = observations/people, columns = variables/questions.


```{r}
# Get the number of rows
nrow(data)

# Get the number of columns
ncol(data)
```

**INTEL**

**What this does**: `nrow()` tells you how many observations you have `ncol()` tells you how many variables were measured.


### What types of data do you have?

```{r}
# See the structure of your dataset
str(data)
```

**INTEL**

**What this does**: This is your data blueprint. Shows you each column name, what type of data it is (numbers, text, etc.), and a preview of the first few values. In other words, it is like getting a research codebook for the study.

```{r}
# Get column names
names(data)
```

**INTEL**

**What this does**: Lists all the column names. Look for patterns - do they follow a naming scheme?

## Phase 2: Looking at Your Data

### Peek at your data

```{r}
# See the first 6 rows
head(data)

# See the first 10 rows
head(data, 10)
```

**INTEL**

**What this does**: Shows you the actual data values. This is like looking at the first page of a survey to see what responses look like.

```{r}
# See the last few rows
tail(data)
```

**INTEL**

**What this does**: Sometimes the end of a dataset is different from the beginning. This checks that.

### Get a statistical overview
```{r}
# Get summary statistics for all columns
summary(data)
```

**INTEL**

**What this does**: For each column, shows you the minimum, maximum, average, median, and other statistics. For text data, it shows the most common values.

## Phase 3: Investigating Specific Columns

### Look at unique values in a column

```{r}
# See all unique values in a specific column (replace "Q1" with any column name)
unique(data$Q1)
```

**INTEL**

**What this does**: Shows you every different value that appears in that column. Great for understanding response scales (like 1-5 ratings) or categories.

```{r}
# Count how many unique values there are
length(unique(data$Q1))
```

**INTEL**

**What this does**: Tells you how many different response options there were. 2 = yes/no, 5 = probably a rating scale, 50+ = maybe open-ended responses.

### Count responses

```{r}
# Count how many times each value appears
table(data$Q1)
```

**INTEL**

**What this does**: Shows you how many people gave each response. Helps you see if most people agreed or if opinions were split.

```{r}
# Get percentages instead of counts
prop.table(table(data$Q1))
```

**INTEL**

**What this does**: Same as above, but shows proportions (0.0 to 1.0) instead of raw numbers.

### Look at missing data
```{r}
# Count missing values in each column
sapply(data, function(x) sum(is.na(x)))
```

**INTEL**

**What this does**: Shows how many people didn't answer each question. High numbers might indicate sensitive questions.



## Pro Tips for Data Detectives

1. **Start big, then zoom in**: Use `dim()` and `str()` first to get the big picture, then use `head()` and `summary()` to see details.

2. **Look for patterns in column names**: Are they Q1, Q2, Q3? Or Age, Income, Education? This tells you about how the data was organized.

3. **Check the range of values**: If you see mostly 1-5 or 1-7, you're probably looking at rating scales. If you see 18-99, might be ages.

4. **Pay attention to missing data**: Questions with lots of missing answers might be sensitive topics.

5. **Use `unique()` liberally**: It's the fastest way to understand what each column actually contains.



## Reference Commands 
### This table will help you when completing your project. Save it!
| Want to know... | Use this code |
|---|---|
| How many rows? | `nrow(data)` |
| How many columns? | `ncol(data)` |
| What does the data look like? | `head(data)` |
| What types of data? | `str(data)` |
| Statistical overview? | `summary(data)` |
| What values are in column Q1? | `unique(data$Q1)` |
| How many people gave each answer to Q1? | `table(data$Q1)` |
| Any missing data? | `sapply(data, function(x) sum(is.na(x)))` |