---
title: "R Code Guide: Bar Plots, Histograms, and Density Plots"
subtitle: "Section 4"
author: "Angelica Remache Lopez"
date: "2025-09-18"
output: pdf_document
---

```{r echo = FALSE}
# Don't run code, but still show chunks when knitting
knitr::opts_chunk$set(eval = FALSE, echo = TRUE)
```


# Introduction

In this section, we are going to learn how to **visualize data** in R using `ggplot2`.  
We’ll start with **bar plots** (categorical data), then move to **histograms** (continuous data), and finally **density plots** (smoothed distributions).  

Remember: **Today is not only about being a good coder, but also about being CREATIVE!**  

You will use these tools to answer questions about immigration perceptions, but first we’ll practice with general examples.

# Load the required packages

```{r}
# Loading the packages we need to be able to produce the graphs
library(dplyr)
library(ggplot2)
```


# Step 1: ggplot Grammar of Graphics

Every plot in `ggplot2` has three parts:

1. **Data**: the dataset you want to use.  
2. **Aesthetics (`aes`)**: how variables are mapped to the plot (x-axis, y-axis, fill, color).  
3. **Geometric layers (`geom_*()`)**: what kind of plot you make (bars, points, densities).  


```{r}
# ggplot skeleton
ggplot(YOUR_DATA, aes(x = XVAR, y = YVAR, fill = GROUP)) +
  geom_something()
```

# Step 2: Bar Plots

Bar plots are best for categorical variables.
We will first plot counts, then plot group means.

```{r}
# Template for counts
YOUR_DATA %>%
  filter(!is.na(CAT_VAR)) %>%
  count(CAT_VAR) %>%
  ggplot(aes(x = reorder(as.factor(CAT_VAR), -n), y = n)) +
  geom_col(fill = "blue") +
  labs(title = "Counts by Category", x = "Category", y = "Count") +
  theme_minimal()
```

This is especially useful for your Assignment 1:

```{r}
# Template for group means
YOUR_DATA %>%
  filter(!is.na(GROUP_VAR), !is.na(RESP_VAR)) %>%
  group_by(GROUP_VAR) %>%
  summarise(mean_resp = mean(RESP_VAR, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(GROUP_VAR, -mean_resp), y = mean_resp)) +
  geom_col(fill = "red") +
  labs(title = "Average Response by Group", x = "Group", y = "Mean Response") +
  theme_classic()

```

# Step 3: Histograms

Histograms show the distribution of a continuous variable.
They work by dividing values into bins.

```{r}
# Histogram template
ggplot(YOUR_DATA, aes(x = NUMERIC_VAR)) +
  geom_histogram(aes(y = ..density..), bins = 20, color = "black", fill = "lightgray") +
  labs(title = "Histogram with Density Scale", x = "Value", y = "Density") +
  theme_minimal()

```

# Step 4: Density Plots

A density plot is a smoothed histogram.
It helps you see the shape of a distribution, and compare groups.

```{r}
# Single density
ggplot(YOUR_DATA, aes(x = NUMERIC_VAR)) +
  geom_density(fill = "steelblue", alpha = 0.4) +
  labs(title = "Density Plot", x = "Value", y = "Density") +
  theme_light()

# By group
ggplot(YOUR_DATA, aes(x = NUMERIC_VAR, fill = GROUP_VAR, color = GROUP_VAR)) +
  geom_density(alpha = 0.25) +
  labs(title = "Density by Group", x = "Value", y = "Density") +
  theme_minimal()

```

