---
title: "Section 3: Worksheet"
subtitle: "Solutions"
author: "Angelica Remache Lopez"
date: "2025-09-09"
output: pdf_document
---

# Worksheet 2: Section 2

## Load your data

```{r}
# Loading data
data <- read.csv("file:///Users/angelica/Desktop/PS138Z_Data/survey_attitudes_immigrants.csv")
```

## What is the total number of observation in the data?

```{r}
# Finding the number of observations/raws 
nrow(data)
```

## What is the total number of variables in the data?

```{r}
# Finding the number of variables/columns
ncol(data)
```

## For how many observations do we have information about wheather they are citizens or have taken any steps to apply for a US citizenship (*S5* in the dataset)?

```{r}
# Getting the total number of responses we have for this questions on US citizenship 
# we are using the symbol ($) because from our data we are choosing only the column/variable "S5"
# which is the question on citizenship

sum(is.na(data$S5))
```
## For how many observations do we have information about wheather they are citizens or have taken any steps to apply for a US citizenship (*S5* in the dataset)?

```{r}
# Getting the total number of N/As we have for this question on US citizenship 
# We use (!) when we reverse our code

sum(!is.na(data$S5))
```


## How many people think that "On the whole, immigration has been a good thing/bad thing/some good/some bad/ don't know for the United States in the past?" (*Q5*)

Note that: 
1= Good thing
2 = Bad thing 
3 = Some good/some bad
9 = Don't know

```{r}

# We can get a table to see the summary of this question and then see the number of respondents for each value
table(data$Q3)

```


